Lens Modeling in \sw

### 2013-09-09 text/tex/spaghetti.tex
edis

@@

Give a guess for the maximum, minimum and saddle points.  Program
tries find $\kappa(x,y)$ that reproduces these properties exactly, and
looks reasonably like a galaxy.  Solution not unique, an ensemble
generated.

\hr

SpaghettiLens is build on top of GLASS \citep{Lubini2012}, that builds and improves upon PixeLens \citep{Saha2004}.
It provides a web based graphical user interface that allows to trace contours with the mouse and identify extremal points by clicking on the image.
\spl then tries to find a mass distribution $\kappa(x,y)$ that reproduces the input parameters.
Since there is no unique solution, \spl samples the solution space and produces an ensemble of solutions, as described in the paper \citep{Lubini2012}.
It provides direct visual feedback by rendering a synthetic image side by side the original image, such that users can directly compare the predictions of their model to the survey image.
It additinally provides plots of the modelled mass distribution and the contour lines of arrival time surface as feedback.

Using this direct visual feedback, even unexperienced users can successfully model lenses using a iterative approach if they know what to look for.

The modelling process consitis of three basic steps:

\begin{enumerate}
  \item identify lensed images and separate them from other background light sources
  \item classify and order images accoring to arrival time. (local minima, maxima, saddlepoints)
  \item fine tune the arrangement, identify addidional external point masses influencing the result
\end{enumerate}

\spl assists volunteers with several features in this process.
Step 1 by supplying several images from several bands (not yet implemented in \sw).
Step 2 by restricting the user, only valid configuration can be entered, the odd number theorem\needcite is taken care of.
Step 3 by providing the visual feedback. Users can check the generated mass distribution and synthetic image.

Explanation:
Synthetic image: A rendering of the derivative of the arrival time for each pixel. This leads to a black and white image that reasonably looks like the visual appearace of the model.

Volunteers enter an assumed lens configuration and get a rendered synthetic image along with a reconstructed / simulated mass map and reconstructed / simulated contour lines.
This allows volunteers to directly see the effect of changing the parameters of the model and to iteratively approach a good solution.

>>>>

In \spl the user input is an educated guess for the topography of the
arrival-time surface.

@@

It provides a web based graphical user interface that allows to trace contours with the mouse and identify extremal points by clicking on the image.

The modelling process consitis of three basic steps:
\begin{enumerate}
  \item identify lensed images and separate them from other background light sources
  \item classify and order images accoring to arrival time. (local minima, maxima, saddlepoints)
  \item fine tune the arrangement, identify addidional external point masses influencing the result
\end{enumerate}

\spl assists volunteers with several features in this process.
Step 1 by supplying several images from several bands (not yet implemented in \sw).
Step 2 by restricting the user, only valid configuration can be entered, the odd number theorem\needcite is taken care of.

Give a guess for the maximum, minimum and saddle points.  Program
tries find $\kappa(x,y)$ that reproduces these properties exactly, and
looks reasonably like a galaxy.  Solution not unique, an ensemble
generated.
SpaghettiLens is build on top of GLASS \citep{Lubini2012}, that builds and improves upon PixeLens \citep{Saha2004}.

\spl then tries to find a mass distribution $\kappa(x,y)$ that reproduces the input parameters.
Since there is no unique solution, \spl samples the solution space and produces an ensemble of solutions, as described in the paper \citep{Lubini2012}.

Step 3 by providing the visual feedback. Users can check the generated mass distribution and synthetic image.

It additinally provides plots of the modelled mass distribution and the contour lines of arrival time surface as feedback.

It provides direct visual feedback by rendering a synthetic image side
by side the original image, such that users can directly compare the
predictions of their model to the survey image. Explanation: Synthetic
image: A rendering of the derivative of the arrival time for each
pixel. This leads to a black and white image that reasonably looks
like the visual appearace of the model.

Using this direct visual feedback, even inexperienced users can
successfully model lenses using a iterative approach if they know what
to look for.
